%-------------------------
% Resume in Latex
% Author : Jake Gutierrez
% Based off of: https://github.com/sb2nov/resume
% License : MIT
%------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage{changepage}  % Provides adjustwidth environment
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}

\usepackage{fontawesome}
\usepackage{multicol}
\setlength{\multicolsep}{-3.0pt}
\setlength{\columnsep}{-1pt}
\usepackage{fontspec}
\usepackage{emoji}
%----------FONT OPTIONS----------
% sans-serif
% \usepackage[sfdefault]{FiraSans}
% \usepackage[sfdefault]{roboto}
% \usepackage[sfdefault]{noto-sans}
% \usepackage[default]{sourcesanspro}

% serif
% \usepackage{CormorantGaramond}
% \usepackage{charter}


\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins
\addtolength{\oddsidemargin}{-0.6in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1.19in}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{1.4in}

\urlstyle{same}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large\bfseries
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

%-------------------------
% Custom commands
\newcommand{\resumeItem}[1]{
  \item\small{
    {#1 \vspace{-2pt}}
  }
}

\newcommand{\classesList}[4]{
    \item\small{
        {#1 #2 #3 #4 \vspace{-2pt}}
  }
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-2pt}\item
    \begin{tabular*}{1.0\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1}  \textit{\small#3} & \textbf{\small #2} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubSubheading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textit{\small#1} & \textit{\small #2} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeProjectHeading}[2]{
    \item
    \begin{tabular*}{1.001\textwidth}{l@{\extracolsep{\fill}}r}
      \small#1 & \textbf{\small #2}\\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubItem}[1]{\resumeItem{#1}\vspace{-4pt}}

\renewcommand\labelitemi{$\vcenter{\hbox{\tiny$\bullet$}}$}
\renewcommand\labelitemii{$\vcenter{\hbox{\tiny$\bullet$}}$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=0.0in, label={}]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

% Patch section to include right-aligned subtitle
\newcommand{\sectionwithnote}[2]{%
  \section[#1]{%
    \noindent
    \makebox[\textwidth]{\textbf{\Large #1} \hfill \textit{#2}}%
  }%
}

%-------------------------------------------
%%%%%%  RESUME STARTS HERE  %%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%----------HEADING----------
% \begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
%   \textbf{\href{http://sourabhbajaj.com/}{\Large Sourabh Bajaj}} & Email : \href{mailto:sourabh@sourabhbajaj.com}{sourabh@sourabhbajaj.com}\\
%   \href{http://sourabhbajaj.com/}{http://www.sourabhbajaj.com} & Mobile : +1-123-456-7890 \\
% \end{tabular*}

\begin{center}
    {\Huge \scshape Paul Charlton} \\ \vspace{4pt}

    \href{mailto:techguru@byiq.com}{\raisebox{-0.2\height}\faEnvelope\ } \quad
    \href{https://github.com/PaulCharlton}{\raisebox{-0.2\height}\faGithub\ } \quad
    \href{https://gitlab.com/techguru}{\raisebox{-0.2\height}\faGitlab\ } \quad
    \href{https://www.linkedin.com/in/techguru/}{\raisebox{-0.2\height}\faLinkedin\ } \quad
    \href{https://paul-charlton.vercel.app/}{\raisebox{-0.2\height}\faGlobe\ } \quad
    \href{https://en.wikipedia.org/wiki/Paul_Charlton_(technologist)}{\raisebox{-0.2\height}\faWikipediaW\ } \quad
    \href{tel:+18082787904}{\raisebox{-0.2\height}\faPhone\ } \quad
    \href{https://signal.me/#p/+18082787904}{\raisebox{-0.2\height}\faComment\ } \quad
    \href{https://wa.me/18082787904}{\raisebox{-0.2\height}\faWhatsapp\ }

    \vspace{-8pt}
\end{center}



%-----------EDUCATION-----------
\sectionwithnote{EDUCATION}{}
  \resumeSubHeadingListStart
    \resumeSubheading
      {Rensselaer Polytechnic Institute}
      {Sept. 1984 -- May 1988}
      {BS: Electrical, Computer, and Systems Engineering}
      {Troy, New York}
  \resumeSubHeadingListEnd

\sectionwithnote{SUMMARY}{Trusted C-Level advisor; Distinguished Engineer with C-Level Impact}
High-trust technical leader operating at the intersection of architecture, innovation, and business strategy. Entrusted with foundational technologies—the corporate “crown jewels”—that generate over \$1T in ARR across multiple Fortune 500 companies, while intentionally remaining a FORCE MULTIPLIER as hands-on IC. Leads critical infrastructure, developer experience, and engineering culture transformation efforts at scale across orgs of 2,000+, and has directly led teams up to 200. \quad Has delivered 100x gains in reliability, latency, and developer productivity through deep systems knowledge, metrics-driven execution, automation of Platform Engineering, and hands-on mentorship. Spearheaded over \$300M+ in investment backing and held pivotal technical roles at Apple (QuickTime) and Intuit (QuickBooks). Drives architecture end-to-end: from IoT device drivers, and desktop applications to cloud-scale SaaS, PaaS, and IaaS—always with a passion for clean systems and ruthless simplicity. \quad Award-winning inventor with multiple patents and a sharp business instinct for IP protection—directly reclaiming over \$150M+ in license revenue for enterprise clients. Communicates fluently with execs and engineers alike; a trusted partner in strategy, innovation, and delivery. Proven driver of stack modernization, reliability engineering, and time-to-value acceleration across complex, regulated environments.

%-----------EXPERIENCE-----------
\sectionwithnote{EXPERIENCE}{curated for relevance to Isomorphic Labs}
  \resumeSubHeadingListStart

\resumeSubheading
    {\href{https://www.virginia.edu/}{University of Virginia - School of Medicine~\faExternalLink}
    \quad\small
    \textit
        {\href {https://paul-charlton.vercel.app/story/9}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {1979 - 1986}
    {Researcher}
    {Charlottesville, VA}
\resumeItem{
    Youngest paid researcher at UVA School of Medicine at 14, recruited after scoring “>170 IQ off the chart” in UVA’s gifted research program.  In a vision research lab under the Department of Physiology, I pioneered the lab-scale integration of analog biomedical equipment with digital computers, automating data collection and experimental workflows in a fully dark lab. I designed and built custom circuits, developed 6502 assembly drivers for GPIB-controlled systems, and implemented an early speaker-independent voice recognition system - advancing published medical research and mentoring graduate-level scientists before entering college.
}

\resumeSubheading
    {\href{https://www.deshaw.com}{D.E. Shaw \& Co.~\faExternalLink}
    \quad\small
    \textit
        {\href {https://paul-charlton.vercel.app/story/25}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {November 1992 -- April 1994}
    {Technology Specialist}
    {New York, NY}
\resumeItem{
    At D. E. Shaw, I joined as employee \#24 and led engineering for Jeff Bezos’s Third Market trading desk, architecting a GUI-based order entry system and scaling infrastructure to support market making in 2,000+ securities. I diagnosed critical system deadlocks in the Stratus OS, optimized predictive trading algorithms using multidimensional statistical calculus, and pioneered in-house logging of historical market data to replace costly external sources. I contributed to early brainstorming sessions that seeded the idea for what became Amazon. My push for employee equity — tying ownership to impact rather than seniority — challenged the status quo and ultimately helped catalyze an exodus that included Jeff’s departure to launch it.
}

\resumeSubheading
    {\href{https://www.apple.com}{Apple Computer Inc.~\faExternalLink}
    \quad\small
    \textit
        {\href{https://paul-charlton.vercel.app/story/11}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {April 1994 -- June 1996}
    {Principal Engineer}
    {Cupertino, CA}
\resumeItem{
    I led the cross-platform port of QuickTime and macOS from proprietary Apple hardware to Windows (x86) by building a fully preemptive, thread-safe implementation of the Macintosh Toolbox from scratch and integrating QuickDraw 2D. I architected and launched the QuickTime Media Layer (QTML), which enabled Mac applications to run on Windows and SGI IRIX. QTML later formed the foundation of Apple’s Carbon API and macOS X transition. My work was pivotal to Apple’s \$150M legal victory over Microsoft/Intel, and I presented the cross-platform QuickTime rollout at WWDC 1996 in the main hall to over 1,000 developers.
}

\resumeSubheading
    {\href{https://www.chase.com}{Chase Manhattan Bank~\faExternalLink}
    \quad\small
    \textit{
        \href{https://paul-charlton.vercel.app/story/17}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {June 1990 -- November 1992}
    {Director of Advanced Technology Lab}
    {New York, NY}
\resumeItem{
    I modernized Chase’s retail and back-office banking systems, replacing paper-based workflows with PC and networked solutions that drastically reduced risk, cut processing time from days to hours, and supported scalable acquisitions. I architected and developed the first ever Windows GUI-based online banking platform, engineered high-security systems for wire transfers and vault reconciliation, and implemented middleware to unify legacy systems through screen-scraping and field validation. My efforts enhanced operational security, uncovered fraud, and saved millions through alternative computing strategies and vendor negotiations.
}

\resumeSubheading
    {\href{https://www.oracle.com/sun}{Sun Microsystems~\faExternalLink}
    \quad\small
    \textit{
        \href{https://paul-charlton.vercel.app/story/4}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {2008}
    {Principal Consulting Engineer}
    {Santa Clara, CA}
\resumeItem{
    In under 10 weeks, I led a 10-person team to rescue and deliver JavaFX 1.0 on time—rewriting the multimedia pipeline, eliminating 30\% of the legacy code, and adding 40,000 lines of high-performance, test-driven code to enable smooth, synchronized video playback across platforms. I re-architected the system for native performance, preemptive multithreading, and customer delight, coordinating closely with product management and external partners. Our release on Dec 4, 2008 marked Sun’s first on-time software launch in over a decade and heralded a 51\% increase in market cap within 15 days.
}
\newpage
\resumeSubheading
    {\href{https://www.oracle.com/sun}{Sun Microsystems - JavaSoft~\faExternalLink}
    \quad\small
    \textit{
        \href{https://paul-charlton.vercel.app/story/5}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {1997}
    {Principal Consulting Engineer}
    {Palo Alto, CA}
\resumeItem{
    As Principal Consulting Engineer, I rearchitected and optimized the Java2D subsystem for JDK 2, boosting glyph rendering performance by 40x to deliver over 200,000 glyphs/sec on 300Mhz Pentium II. I also applied low-level Intel i386 pipelining techniques to the Java Bytecode interpreter, achieving a 5x performance gain. This foundational code remains virtually unchanged in every Android device today and across graphical Java-based systems globally.
}

\resumeSubheading
    {\href{https://www.hp.com}{Hewlett-Packard~\faExternalLink}
    \quad\small
    \textit{
        \href{https://paul-charlton.vercel.app/story/38}{[Full Story]~{\fontspec{Symbola}\symbol{"1F517}}}}
    }
    {June 1988 -- May 1990}
    {Hardware and Software Engineer}
    {Greeley, CO}
\resumeItem{
    At Hewlett-Packard, I developed advanced SCSI-II drivers and boot ROMs for magneto-optic and tape storage devices across HP-UX, MPE, and Apollo Domain OS, including debugging low-level OS synchronization flaws and pioneering in-house ASIC/VLSI chip design with a successful first tape-out. I led a data recovery initiative for NASA's historic space mission telemetry, creating custom software and decoding pipelines for degraded tape reels, and contributed to integrating HP storage with Apollo workstations post-acquisition. Certified as a Six Sigma coach, I worked cross-functionally to enforce defect prevention at the requirements stage and applied HP's culture of merit-driven, engineer-led innovation.
}

\resumeSubHeadingListEnd
\vspace{-16pt}

\section*{Publications, Patents and Products}

\textbf{Key Products:} Apple QuickTime (iTunes, iOS), Apple Carbon API, JavaFX, Java2D (Android), MDOS, QuickBooks

\vspace{0.5em}
\textbf{Selected Publications:}
\noindent
Morton, R. W., Chung, J. K., Miller, J. L., Charlton, J. P., \& Fager, R. S. (1986). \textit{Extended sensitivity for the calcium selective electrode}. Analytical Biochemistry, 157(2), 345–352. \href{https://scholar.google.com/scholar_lookup?title=Extended+sensitivity+for+the+calcium+selective+electrode&author=Morton&publication_year=1986&journal=Analytical+Biochemistry&volume=157&pages=345-352&doi=10.1016/0003-2697(86)90636-6}\newline
{Google Scholar~\faExternalLink}.  \href{https://doi.org/10.1016/0003-2697(86)90636-6}{DOI: 10.1016/0003-2697(86)90636-6}

\vspace{0.5em}
\textbf{Selected Patents:}

\textbf{\href{https://patents.google.com/patent/US5875354}{US 5,875,354~\faExternalLink}} (1999) – \textit{System for Synchronization by Modifying the Rate of Conversion by Difference of Rate Between First Clock and Audio Clock}
\textbullet{} Inventors: Paul Charlton, Keith Gurganus; Assignee: Apple Inc.; Early solution to audio/video sync via dynamic sample rate conversion. Now core to AV streaming, DAWs, AI video rendering, and IoT synchronization.

\vspace{0.5em}

\textbf{\href{https://patents.google.com/patent/US5825359}{US 5,825,359~\faExternalLink}} (1998) – \textit{Improved Arbitration of a Display Screen in a Computer System}
\textbullet{} Queue-based screen access control in multi-process environments.; Anticipated today’s GPU compositing (Quartz, DWM, Wayland) and virtual desktop rendering.

\vspace{0.5em}

\textbf{\href{https://patents.google.com/patent/US5949434}{US 5,949,434~\faExternalLink}} (1999) – \textit{Scaling 2D Graphic Images Without Banding Artifacts}
\textbullet{} Enhances Bresenham’s algorithm with pseudo-random shifts to reduce visual artifacts.; Forefather of adaptive upscaling in DLSS, medical imaging, and AI-based super-resolution.

%-----------PROJECTS-----------
% \section{Projects}
%     \vspace{-5pt}
%     \resumeSubHeadingListStart
%       \resumeProjectHeading
%           {\textbf{Gym Reservation Bot} $|$ \emph{Python, Selenium, Google Cloud Console}}{January 2021}
%           \resumeItemListStart
%             \resumeItem{Developed an automatic bot using Python and Google Cloud Console to register myself for a timeslot at my school gym.}
%             \resumeItem{Implemented Selenium to create an instance of Chrome in order to interact with the correct elements of the web page.}
%             \resumeItem{Created a Linux virtual machine to run on Google Cloud so that the program is able to run everyday from the cloud.}
%             \resumeItem{Used Cron to schedule the program to execute automatically at 11 AM every morning so a reservation is made for me.}
%           \resumeItemListEnd
%           \vspace{-13pt}
%       \resumeProjectHeading
%           {\textbf{Ticket Price Calculator App} $|$ \emph{Java, Android Studio}}{November 2020}
%           \resumeItemListStart
%             \resumeItem{Created an Android application using Java and Android Studio to calculate ticket prices for trips to museums in NYC.}
%             \resumeItem{Processed user inputted information in the back-end of the app to return a subtotal price based on the tickets selected.}
%             \resumeItem{Utilized the layout editor to create a UI for the application in order to allow different scenes to interact with each other.}
%           \resumeItemListEnd
%           \vspace{-13pt}
%           \resumeProjectHeading
%           {\textbf{Transaction Management GUI} $|$ \emph{Java, Eclipse, JavaFX}}{October 2020}
%           \resumeItemListStart
%             \resumeItem{Designed a sample banking transaction system using Java to simulate the common functions of using a bank account.}
%             \resumeItem{Used JavaFX to create a GUI that supports actions such as creating an account, deposit, withdraw, list all acounts, etc.}
%             \resumeItem{Implemented object-oriented programming practices such as inheritance to create different account types and databases.}
%           \resumeItemListEnd
%     \resumeSubHeadingListEnd
% \vspace{-15pt}


%
%-----------PROGRAMMING SKILLS-----------
\sectionwithnote{Technologies \& Keywords}{Machine Readable}
\fontsize{3pt}{4pt}\selectfont{
\begin{itemize}[leftmargin=0.1in, label={}, itemsep=0pt, parsep=0pt, topsep=0pt, partopsep=0pt]
\item
    \textbf{\underline{Hardware Interfaces:}} \hspace{0pt}
    AGP - Accelerated Graphics Port;
    Centronics - Centronics Parallel Interface;
    EISA - Extended Industry Standard Architecture;
    eSATA - External Serial ATA;
    FireWire - IEEE 1394 High Performance Serial Bus;
    GPIB - General Purpose Interface Bus;
    HDMI - High-Definition Multimedia Interface;
    HPIB - Hewlett-Packard Interface Bus;
    ISA - Industry Standard Architecture;
    NuBus - New Bus Architecture;
    PATA - Parallel Advanced Technology Attachment;
    PCI - Peripheral Component Interconnect;
    PCIe - Peripheral Component Interconnect Express;
    PCMCIA - Personal Computer Memory Card International Association;
    RS-232 - serial interface;
    RS-422 - serial interface;
    RS-485 - serial interface;
    S-100 - S-100 Bus;
    SAS - Serial Attached SCSI;
    SATA - Serial Advanced Technology Attachment;
    SCSI-II - Small Computer System Interface II;
    ST506 - Seagate Technology ST-506 Interface;
    Thunderbolt - Thunderbolt Interface;
    USB - Universal Serial Bus 1.x, 2.x;
    VLB - VESA Local Bus

\item
    \textbf{\underline{Processors \& Architectures:}} \hspace{0pt}
    6502;
    6809;
    8088;
    Alpha (DEC Alpha);
    AMD29K;
    AMD bit-slice;
    ARM (ARMv4–v9);
    IBM 360;
    Itanium (IA-64);
    M68000;
    MIPS;
    PA-RISC;
    PDP-11;
    PowerPC (PPC);
    RCA COSMAC 1802;
    RISC-V;
    RS6000;
    SPARC;
    TMS9900;
    VAX;
    Z80;
    x86 (16-bit 8086);
    x86\_32 (IA-32);
    x86\_64 (AMD64)

\item
    \textbf{\underline{Systems \& Platforms:}} \hspace{0pt}
    AIX;
    AmigaOS;
    Android;
    Apollo Domain OS;
    BSD (FreeBSD, OpenBSD, NetBSD);
    CP/M;
    DR-DOS;
    FreeRTOS;
    HPUX;
    IBM 360;
    iOS;
    Linux (kernel and distros);
    macOS;
    MINIX;
    MPE;
    MS-DOS;
    Multics;
    MTS;
    MVS;
    MyArc Geneve;
    NDOS;
    OS/2;
    QNX;
    RTOS;
    SGI IRIX;
    SGI ONYX;
    Solaris;
    TI-99/4A;
    Unix;
    VOS;
    VxWorks;
    Windows 3.0;
    Windows 3.1;
    Windows NT

\item
    \textbf{\underline{Drivers \& Low-level Development:}} \hspace{0pt}
    Device Tree Overlays (Linux ARM platforms);
    HPUX drivers (68020, C);
    Linux kernel modules (loadable drivers, LKM);
    Linux VFS (Virtual Filesystem Switch);
    MPE drivers (Pascal);
    MS-DOS INT 13h BIOS-level block device handlers;
    macOS drivers;
    macOS I/O Kit (C++/Objective-C);
    low-level filesystem drivers;
    low-level filesystems - XFS and Clustered-XFS;
    PCI configuration space access;
    RSX-11 and VAX/VMS device drivers (MACRO-11, BLISS);
    SCSI-II Autochanger drivers;
    SCSI-II Magneto-optic drivers;
    Solaris Device Driver Interface (DDI/DKI);
    TI-99/4A CRU (Communications Register Unit) drivers (assembly);
    Windows 3.x DDK;
    Windows DDK (Device Driver Kit);
    Windows NT DDK;
    BIOS bootloader development;
    EFI (Extensible Firmware Interface) programming;
    GRUB bootloader configuration and customization;
    Secure Boot chain validation and integration.

\item
    \textbf{\underline{Programming Languages \& Tools:}}  \hspace{0pt}
    .NET;
    6502 assembly;
    6809 assembly;
    8088 assembly;
    Alpha assembly;
    AMD bit-slice;
    AMD29K assembly;
    Apollo DSEE (Domain Software Engineering Environment);
    ARM assembly;
    bash;
    bison;
    Boost (C++ Libraries);
    Borland C++;
    BoundsChecker;
    C;
    C\#;
    C++;
    C++ template metaprogramming;
    Carbon;
    ClearCase;
    CodeWarrior;
    Cucumber;
    CVS;
    Eclipse (IDE);
    Enterprise Architect (Sparx Systems);
    flex;
    gcc;
    gdb;
    gem;
    Git (Version Control);
    Go (Golang);
    gprof;
    HyperCard;
    IDL (Interface Definition Language);
    IBM 360 assembly;
    IntelliJ IDEA (IDE);
    Itanium assembly;
    Java;
    JavaScript (browser);
    JDK (Java Development Kit);
    JetBrains MPS;
    JNI (Java Native Interface);
    JProfiler;
    JUnit;
    ld;
    lex;
    Linux kernel modules;
    lldb;
    M68000 assembly;
    macOS I/O Kit;
    macOS SDK;
    make;
    MIPS assembly;
    Modern shell (sh, bash, zsh);
    MS-DOS INT 13h handlers;
    NetBeans (IDE);
    Node.js;
    npm;
    nvm;
    NUnit;
    OpenAPI (formerly Swagger);
    OpenGL;
    PA-RISC assembly;
    Pascal;
    Perforce (Version Control);
    Perl;
    PDP-11 assembly;
    Postman;
    PowerPC assembly;
    Process Explorer;
    Procmon;
    Protocol Buffers (protobuf);
    PsExec;
    Purify;
    QTML;
    QuickDraw;
    QuickTime SDK;
    RCS (Version Control);
    RCA COSMAC 1802 assembly;
    React;
    RISC-V assembly;
    RSpec;
    RS6000 assembly;
    Ruby;
    Salad;
    SCCS (Version Control);
    sed;
    Selenium;
    shell (sh, csh);
    SoftICE;
    SPARC assembly;
    Subversion (SVN Version Control);
    Swagger (OpenAPI predecessor);
    Sysinternals Suite;
    TI GPL (Graphics Programming Language);
    TMS9900 assembly;
    TypeScript;
    UML (Unified Modeling Language);
    Unix utilities (awk, sed, grep);
    Valgrind;
    VAX assembly;
    vi;
    vim;
    Visual Studio (IDE);
    VisualVM;
    VTune;
    webpack;
    Win32 API;
    Windows 3.x DDK;
    Windows DDK;
    Windows NT DDK;
    Windows SDK;
    Windbg;
    X11;
    Xcode;
    XNU kernel;
    Xt;
    x86 (16-bit 8086) assembly;
    x86\_32 (IA-32) assembly;
    x86\_64 (AMD64) assembly;
    yacc;
    yarn;
    Z80 assembly
\item
    \textbf{\underline{UI \& Graphics:}} \hspace{0pt}
    AWT (Abstract Window Toolkit);
    Cross-Platform QuickTime;
    Direct2D;
    Fonts and Typography;
    Geometry;
    Glyphs;
    GTK+;
    HTML5 Canvas;
    Java2D;
    JavaFX;
    macOS UI;
    Microsoft GDI (Graphics Device Interface);
    OpenGL;
    Qt;
    Quartz 2D;
    QuickDraw 2D;
    QuickTime 1.x--3.x;
    SVG;
    Swing;
    Windows GUI;
    X11;
\item
    \textbf{\underline{HID: Graphics, Video\& Sound:}} \hspace{0pt}
    DirectX;
    DisplayPort;
    dropping frames;
    FFmpeg;
    graphics performance;
    H.264;
    HDMI;
    jitter;
    latency;
    NVENC;
    OpenGL;
    sound card drivers;
    VGA drivers;
    Video Capture;
    Video Capture Card Drivers;
    Video Capture Cards;
    Video Compression;
    Video Decompression;
    video pipeline;
    video transcoding;
    WebRTC;
    YUV;
\item
    \textbf{\underline{Statistical \& Optimization Methods:}} \hspace{0pt}
    6-sigma;
    A/B testing;
    gradient ascent;
    gradient boosting;
    gradient descent;
    K-means clustering;
    linear regression;
    logistic regression;
    monte carlo simulation;
    Multidimensional linear regression;
    simulated annealing;
    stochastic gradient descent;
    SVD (Singular Value Decomposition);
\item
    \textbf{\underline{Networking \& Protocols:}} \hspace{0pt}
    802.11;
    BGP;
    Bluetooth;
    Bluetooth Low Energy;
    CAN bus;
    DNS;
    Ethernet;
    FTP;
    HTTP;
    HTTPS;
    IMAP;
    IPv4;
    IPv6;
    ISDN (Integrated Services Digital Network);
    leased lines;
    LoRaWAN;
    LU2;
    LU6.2;
    Modbus;
    MQTT;
    Novell NetWare;
    POP3;
    SMTP;
    SNA (System Network Architecture);
    SNMP;
    SSH;
    T1;
    TCP/IP;
    Telnet;
    TLS;
    Token Ring;
    UDP;
    WebSocket;
    Wi-Fi;

\item
    \textbf{\underline{Enterprise \& Legacy Systems:}} \hspace{0pt}
    AS/400;
    CICS (Customer Information Control System);
    COBOL;
    DB2;
    HP 3000;
    JCL (Job Control Language);
    MPE;
    MTS (Michigan Terminal System);
    MVS (Multiple Virtual Storage);
    RSTS;
    RSX-11;
    System/360;
    System/370;
    TPF (Transaction Processing Facility);
    VAX/VMS;

\item
    \textbf{\underline{Project \& Product Experience:}} \hspace{0pt}
    Apollo Computer Acquisition;
    Apple StarTrek;
    Beckman;
    BeyondNews;
    Bio-Rad;
    D. E. Shaw;
    EG\&G;
    Element Analytics;
    F5 Networks;
    Fisher Scientific;
    Forticom;
    GIMPS;
    Hewlett Packard;
    IdeaLab;
    Independa;
    Intacct;
    Intuit;
    NASA;
    NASDAQ;
    National Merit;
    Netflix Prize;
    Neustar;
    NYSE;
    Perkin-Elmer;
    Pixar;
    Tektronix;
    U.S. Coast Guard Auxiliary;
    ValueClick;
    Varian;
    Waters HPLC;

\item
    \textbf{\underline{OS Internals \& Concepts:}} \hspace{0pt}
    Async distributed software;
    bootloaders;
    caching
    device driver abstraction;
    file descriptor management;
    Global-scale time synchronization algorithms;
    init systems;
    interrupt handlers;
    kernel scheduling;
    memory management;
    paging;
    preemption;
    process isolation;
    resource management;
    semaphores;
    system calls;
    virtual memory management;

\item
    \textbf{\underline{Development Methodologies:}} \hspace{0pt}
    Agile;
    Behavior-Driven Development (BDD);
    Documentation-driven design;
    Kanban;
    long locks held across autochanger seeks;
    Model-Driven Development (MDD);
    product management;
    Requirements before Design;
    Scrum;
    Test-Driven Development (TDD);
    Waterfall;

\item
    \textbf{\underline{Vintage Computing:}} \hspace{0pt}
    5" Floppy (FM/MFM formatting);
    8" Floppy (MFM formatting);
    Altair 8800;
    Apple II;
    Apple III;
    Commodore Amiga;
    Commodore PET;
    Commodore SuperPET;
    CP/M;
    CRT terminals;
    DEC PDP-11;
    IBM PC XT;
    KayPro;
    paper tape readers;
    S100 bus;
    Seequa Chameleon;
    Teletypes;
    TI-99/4A;
    TRS-80;

\item
    \textbf{\underline{Language Support \& Internationalization:}} \hspace{0pt}
    bidirectional text;
    CJK (Chinese, Japanese, Korean) language support;
    codepage translation;
    collation algorithms;
    date and time formatting;
    font fallback;
    ICU (International Components for Unicode);
    ICU collation services;
    input method editors (IME);
    language negotiation;
    locale-aware string sorting;
    locale handling;
    LTR and RTL layout handling;
    multi-script rendering;
    natural language collation;
    number and currency formatting;
    pluralization rules;
    sort key generation;
    string normalization;
    surrogate pairs;
    text segmentation;
    transliteration;
    Unicode code points;
    Unicode Internals;
    Unicode normalization forms (NFC, NFD, NFKC, NFKD);
    Unicode scripts;
    UTF-16;
    UTF-8;
    variant selectors;
    zero-width joiners;

\item
    \textbf{\underline{Human-Computer Interaction:}} \hspace{0pt}
    command-line interfaces;
    dialog systems;
    eye tracking;
    multimodal interfaces;
    Speaker-independent speech recognition;
    Speech synthesis;
    touchscreen UI;
\item
    \textbf{\underline{Cloud Scale Platform Engineering:}} \hspace{0pt}
    ACME protocol (TLS automation);
    Apache Kafka;
    Apache Spark;
    API backward compatibility;
    Attribute-based access control (ABAC);
    Blue/green deployments;
    Canary deployments;
    Centralized log aggregation;
    Centralized policy engine (e.g., OPA);
    Ceph;
    CI/CD enforcement with SAST, DAST, TDD;
    Cloud-init;
    Cloudflare (CDN);
    Code coverage gating;
    Consul;
    Container image SemVer tagging;
    Containerd;
    CoreDNS;
    CVE scanning policy;
    Declarative orchestration (e.g., Kubernetes);
    Distributed tracing (OpenTelemetry);
    Docker;
    ElasticSearch;
    etcd;
    Feature flag routing;
    Federated authentication (OIDC/SAML);
    Fine-grained RBAC;
    Flannel;
    Git commit SHA tagging;
    GitOps change management;
    gRPC;
    Helm;
    Horizontal autoscaling policies;
    Hyperscalers (AWS, GCP, AZURE);
    Immutable audit trails;
    Integration test harness;
    Istio;
    JWT or SPIFFE/SPIRE identity;
    Kubernetes (K8s);
    Linkerd;
    Load and chaos testing;
    Log redaction and privacy rules;
    Logging access controls;
    Logging format: structured JSON;
    Metric ontologies (RED, USE);
    MinIO;
    mTLS enforcement (Istio, Linkerd);
    Multi-zone / HA deployment policy;
    NGINX;
    One process per container;
    Open Policy Agent (OPA);
    OpenTelemetry;
    Policy versioning and rollback;
    Production-like staging environments;
    Prometheus;
    Pulumi;
    Redis;
    Regression test suite on merge;
    Resource quotas and limits;
    S3;
    SBOM (Software Bill of Materials);
    Schema migration strategy;
    Secrets management;
    Secure Development Lifecycle (SDL);
    Secure key vault integration;
    Service mesh enforcement;
    Session management policies;
    Sharding;
    SLA-based alerting and load shaping;
    SLO dashboards;
    SOC-2 Type II compliance;
    Structured audit logging;
    Synthetic functional tests in production;
    Terraform;
    TLS cipher suite and protocol policy;
    Token lifecycle and revocation;
    Trace propagation and sampling;
    Vault (HashiCorp);
    Versioned architecture standards;
    Zero-downtime deployments;
    Zookeeper;

\end{itemize}
}

% %-----------INVOLVEMENT---------------
% \section{Leadership / Extracurricular}
%     \resumeSubHeadingListStart
%         \resumeSubheading{Fraternity}{Spring 2020 -- Present}{President}{University Name}
%             \resumeItemListStart
%                 \resumeItem{Achieved a 4 star fraternity ranking by the Office of Fraternity and Sorority Affairs (highest possible ranking).}
%                 \resumeItem{Managed executive board of 5 members and ran weekly meetings to oversee progress in essential parts of the chapter.}
%                 \resumeItem{Led chapter of 30+ members to work towards goals that improve and promote community service, academics, and unity.}
%             \resumeItemListEnd

%     \resumeSubHeadingListEnd


\end{document}
