%-------------------------
% Resume in Latex
% Author : Paul Charlton (based on Jake Gutierrez template)
%------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage{changepage}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}

\usepackage{fontawesome5}
% --- begin: FA4→FA5 external-link shim ---
\makeatletter
\@ifundefined{faExternalLink}{
    \newcommand{\faExternalLink}{\faIcon{external-link-alt}}
}{
    \renewcommand{\faExternalLink}{\faIcon{external-link-alt}}
}
\makeatother
% --- end: FA4→FA5 external-link shim ---

\usepackage{multicol}
\setlength{\multicolsep}{-3.0pt}
\setlength{\columnsep}{-1pt}
\usepackage{fontspec}
\usepackage{emoji}
\usepackage[normalem]{ulem}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\addtolength{\oddsidemargin}{-0.6in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1.19in}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{1.4in}

\urlstyle{same}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

\titleformat{\section}{
    \vspace{-4pt}\scshape\raggedright\large\bfseries
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

% Custom commands
\newcommand{\resumeItem}[1]{\item\small{{#1 \vspace{-2pt}}}}
\newcommand{\resumeSubheading}[4]{
    \vspace{-2pt}\item
    \begin{tabular*}{1.0\textwidth}[t]{l@{\extracolsep{\fill}}r}
    \textbf{#1}  \textit{\small#3} & \textbf{\small #2} \\
    \end{tabular*}\vspace{-7pt}
}
\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=0.0in, label={}]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

\IfFontExistsTF{Symbola}{
    \newfontfamily\LinkFont{Symbola}[Scale=MatchLowercase]
}{
    \IfFontExistsTF{Noto Emoji}{
        \newfontfamily\LinkFont{Noto Emoji}[Scale=MatchLowercase]
    }{
        \newfontfamily\LinkFont{Unifont}[Scale=MatchLowercase] % optional fallback
    }
}

\newcommand{\linkglyph}{{\LinkFont\symbol{"1F517}}}

% Patch section to include right-aligned subtitle
\newcommand{\sectionwithnote}[2]{%
    \section[#1]{%
        \noindent
        \makebox[\textwidth]{\textbf{\Large #1} \hfill \textit{#2}}%
    }%
}

%-------------------------------------------
\begin{document}

%----------HEADING----------
    \begin{center}
    {\Huge \scshape Paul Charlton} \\
    \vspace{4pt}
    \href{mailto:techguru@byiq.com}{\faEnvelope} \quad
    \href{https://github.com/PaulCharlton}{\faGithub} \quad
    \href{https://gitlab.com/techguru}{\faGitlab} \quad
    \href{https://www.linkedin.com/in/techguru/}{\faLinkedin} \quad
    \href{https://paul-charlton.vercel.app/}{\faGlobe} \quad
    \href{https://en.wikipedia.org/wiki/Paul_Charlton_(technologist)}{\faWikipediaW} \quad
    \href{tel:+18082787904}{\faPhone} \quad
    \href{https://signal.me/#p/+18082787904}{\faComment} \quad
    \href{https://wa.me/18082787904}{\faWhatsapp}
    \vspace{-8pt}
    \end{center}

%-----------EDUCATION-----------
    \sectionwithnote{EDUCATION}{}
    \resumeSubHeadingListStart
    \resumeSubheading
    {Rensselaer Polytechnic Institute}
    {Sept. 1984 -- May 1988}
    {BS: Electrical, Computer, and Systems Engineering}
    {Troy, NY}
    \resumeSubHeadingListEnd

%-----------SUMMARY-----------
    \sectionwithnote{SUMMARY}{Principal Engineer \& Architect – Last Decade Focus}
    \begin{itemize}
        \item Distinguished Engineer operating as a high-leverage IC with consistent C-level impact. In the past decade: stabilized clustered filesystems at SGI/HPE, built IoT edge-to-cloud platforms at Element Analytics, scaled Acuity Brands’ IoT cloud to 100B+ events/day, advised Rocky Linux on kernel/SBOM lineage, and founded ByIQ venture studio for sovereign OS and secure comms.
        \item Recognized mentor and multiplier: led org-wide transformations in companies of 600+ engineers; coached teams through cultural and technical migrations; praised as a “walking MSDN” and cultural change agent.
        \item Trusted by Fortune 100 and open-source communities alike; known for ruthless simplicity in systems design, while anchoring technical decisions in business outcomes.
    \end{itemize}

%-----------EXPERIENCE-----------
    \sectionwithnote{EXPERIENCE}{2014–Present Only}
    \resumeSubHeadingListStart

    \resumeSubheading
    {SGI / HPE}{2014 -- 2017}
    {Consulting Architect \& Principal Engineer}{}
    \resumeItemListStart
    \resumeItem{Stabilized and modernized CXFS clustered filesystem with direct C-level revenue impact.}
    \resumeItem{Diagnosed kernel/filesystem synchronization issues; delivered systemic fixes protecting key enterprise accounts.}
    \resumeItemListEnd

    \resumeSubheading
    {Element Analytics}{Dec 2017 -- Aug 2019}
    {Principal Architect (IoT Edge-to-Cloud)}{}
    \resumeItemListStart
    \resumeItem{Designed Kubernetes clusters with GitOps, Helm, IaC, and semver pipelines.}
    \resumeItem{Unified refinery sensor data, process diagrams, and maintenance logs into normalized time-series models for anomaly detection.}
    \resumeItem{Embedded DevSecOps into Azure IoT pipelines; coached operators and engineers at BP, Shell, and Chevron.}
    \resumeItemListEnd

    \resumeSubheading
    {Acuity Brands (Fortune 1000)}{Sept 2019 -- Feb 2022}
    {Chief Cloud Officer}{}
    \resumeItemListStart
    \resumeItem{Directed cloud/IoT strategy across 600 engineers and 5 divisions; scaled platform to 100B+ events/day.}
    \resumeItem{Built SOC2-compliant CI/CD with commit-to-prod under 20 minutes; integrated CVE scanning.}
    \resumeItem{Guided migration of 800+ .NET/Cobol services into microservices; coached teams through cultural transformation.}
    \resumeItem{Partnered with Walmart, Home Depot, CVS, Walgreens to deploy BLE mesh IoT enabling 30cm indoor positioning.}
    \resumeItemListEnd

    \resumeSubheading
    {CIQ / Rocky Linux}{Dec 2021 -- Jul 2023}
    {Board of Advisors}{}
    \resumeItemListStart
    \resumeItem{Led Singularity $\rightarrow$ Apptainer port, abstracting proprietary features for community continuity.}
    \resumeItem{Supported Tier 3 kernel escalations (Rakuten RAN, NVIDIA drivers, kernel race conditions).}
    \resumeItem{Architected SBOM system ingesting kernels, distros, userland; queryable by binary for CVE lineage.}
    \resumeItem{Mentored open-source contributors while upstreaming RT kernel patches.}
    \resumeItemListEnd

    \resumeSubheading
    {Results ByIQ LLC}{Jul 2023 -- Present}
    {Founder \& Principal Researcher}{}
    \resumeItemListStart
    \resumeItem{Founded venture studio exploring sovereign OS, reproducible builds, and secure comms.}
    \resumeItem{Incubated projects such as Ambient OS (encrypted memory, programmable consent), DataGator (forensic SBOM lineage), Forticom (defense-grade comms).}
    \resumeItem{Actively advising startups while remaining fully IC-focused.}
    \resumeItemListEnd

    \resumeSubHeadingListEnd

    \sectionwithnote{Earlier Career}{Pre-2014}
    \begin{itemize}
        \item Principal engineer/architect roles at Apple (cross-platform QuickTime), Sun (Java2D/JavaFX), Intuit (QuickBooks), Cisco (IOS multicore refactor), HP (storage drivers), DE Shaw (trading infra). Trusted as mentor and change agent across engineering cultures.
    \end{itemize}

%-----------TECHNOLOGIES-----------
    \section*{Publications, Patents and Products}

    \textbf{Key Products underpinning multiple \$1T+ industries}: QuickTime (basis for MP4/streaming), Carbon API (bridge to OS X/iOS), Java2D/JavaFX (graphics/UI pipelines), QuickBooks (SMB fintech rails), MDOS (GUI/VM OS), Chase OLB (first GUI home banking).  DE Shaw, pioneered quant arbitrage models using SVD, gradient descent, and simulated annealing feedback loops, foreshadowing modern AI/ML optimization.

    \vspace{0.5em}
    \textbf{Selected Publications:}
    \noindent
    Morton, R. W., Chung, J. K., Miller, J. L., Charlton, J. P., \& Fager, R. S. (1986). \textit{Extended sensitivity for the calcium selective electrode}. Analytical Biochemistry, 157(2), 345–352.\newline
    \href{https://scholar.google.com/scholar_lookup?title=Extended+sensitivity+for+the+calcium+selective+electrode&author=Morton&publication_year=1986&journal=Analytical+Biochemistry&volume=157&pages=345-352&doi=10.1016/0003-2697(86)90636-6}{Google Scholar~\faExternalLink}
    \href{https://doi.org/10.1016/0003-2697(86)90636-6}{DOI: 10.1016/0003-2697(86)90636-6}

    \vspace{0.5em}
    \textbf{Selected Patents:}

    \textbf{\href{https://patents.google.com/patent/US5875354}{US 5,875,354~\faExternalLink}} – \textit{System for Synchronization by Modifying the Rate of Conversion by Difference of Rate Between First Clock and Audio Clock}
    \textbullet{} Inventors: Paul Charlton, Keith Gurganus; Assignee: Apple Inc.; Early solution to audio/video sync via dynamic sample rate conversion. Now core to AV streaming, DAWs, AI video rendering, and IoT synchronization.

    \vspace{0.5em}

    \textbf{\href{https://patents.google.com/patent/US5825359}{US 5,825,359~\faExternalLink}} – \textit{Improved Arbitration of a Display Screen in a Computer System}
    \textbullet{} Queue-based screen access control in multi-process environments.; Anticipated today’s GPU compositing (Quartz, DWM, Wayland) and virtual desktop rendering.

    \vspace{0.5em}

    \textbf{\href{https://patents.google.com/patent/US5949434}{US 5,949,434~\faExternalLink}} – \textit{Scaling 2D Graphic Images Without Banding Artifacts}
    \textbullet{} Enhances Bresenham’s algorithm with pseudo-random shifts to reduce visual artifacts.; Forefather of adaptive upscaling in DLSS, medical imaging, and AI-based super-resolution.

%-----------PROJECTS-----------
% \section{Projects}
%     \vspace{-5pt}
%     \resumeSubHeadingListStart
%       \resumeProjectHeading
%           {\textbf{Gym Reservation Bot} $|$ \emph{Python, Selenium, Google Cloud Console}}{January 2021}
%           \resumeItemListStart
%             \resumeItem{Developed an automatic bot using Python and Google Cloud Console to register myself for a timeslot at my school gym.}
%             \resumeItem{Implemented Selenium to create an instance of Chrome in order to interact with the correct elements of the web page.}
%             \resumeItem{Created a Linux virtual machine to run on Google Cloud so that the program is able to run everyday from the cloud.}
%             \resumeItem{Used Cron to schedule the program to execute automatically at 11 AM every morning so a reservation is made for me.}
%           \resumeItemListEnd
%           \vspace{-13pt}
%       \resumeProjectHeading
%           {\textbf{Ticket Price Calculator App} $|$ \emph{Java, Android Studio}}{November 2020}
%           \resumeItemListStart
%             \resumeItem{Created an Android application using Java and Android Studio to calculate ticket prices for trips to museums in NYC.}
%             \resumeItem{Processed user inputted information in the back-end of the app to return a subtotal price based on the tickets selected.}
%             \resumeItem{Utilized the layout editor to create a UI for the application in order to allow different scenes to interact with each other.}
%           \resumeItemListEnd
%           \vspace{-13pt}
%           \resumeProjectHeading
%           {\textbf{Transaction Management GUI} $|$ \emph{Java, Eclipse, JavaFX}}{October 2020}
%           \resumeItemListStart
%             \resumeItem{Designed a sample banking transaction system using Java to simulate the common functions of using a bank account.}
%             \resumeItem{Used JavaFX to create a GUI that supports actions such as creating an account, deposit, withdraw, list all acounts, etc.}
%             \resumeItem{Implemented object-oriented programming practices such as inheritance to create different account types and databases.}
%           \resumeItemListEnd
%     \resumeSubHeadingListEnd
% \vspace{-15pt}


%
%-----------PROGRAMMING SKILLS-----------
    \sectionwithnote{Technologies \& Keywords}{Machine Readable}
    \begin{multicols}{3}
        \fontsize{2pt}{3pt}\selectfont{
            \begin{itemize}[leftmargin=0.1in, label={}, itemsep=0pt, parsep=0pt, topsep=0pt, partopsep=0pt]
                \item
                \textbf{\uline{Hardware Interfaces:}} \hspace{0pt}
                AGP - Accelerated Graphics Port;
                Centronics - Centronics Parallel Interface;
                EISA - Extended Industry Standard Architecture;
                eSATA - External Serial ATA;
                FireWire - IEEE 1394 High Performance Serial Bus;
                GPIB - General Purpose Interface Bus;
                HDMI - High-Definition Multimedia Interface;
                HPIB - Hewlett-Packard Interface Bus;
                ISA - Industry Standard Architecture;
                NuBus - New Bus Architecture;
                PATA - Parallel Advanced Technology Attachment;
                PCI - Peripheral Component Interconnect;
                PCIe - Peripheral Component Interconnect Express;
                PCMCIA - Personal Computer Memory Card International Association;
                RS-232 - serial interface;
                RS-422 - serial interface;
                RS-485 - serial interface;
                S-100 - S-100 Bus;
                SAS - Serial Attached SCSI;
                SATA - Serial Advanced Technology Attachment;
                SCSI-II - Small Computer System Interface II;
                ST506 - Seagate Technology ST-506 Interface;
                Thunderbolt - Thunderbolt Interface;
                USB - Universal Serial Bus 1.x, 2.x;
                VLB - VESA Local Bus

                \item
                \textbf{\uline{Processors \& Architectures:}} \hspace{0pt}
                6502;
                6809;
                8088;
                Alpha (DEC Alpha);
                AMD29K;
                AMD bit-slice;
                ARM (ARMv4–v9);
                IBM 360;
                Itanium (IA-64);
                M68000;
                MIPS;
                PA-RISC;
                PDP-11;
                PowerPC (PPC);
                RCA COSMAC 1802;
                RISC-V;
                RS6000;
                SPARC;
                TMS9900;
                VAX;
                Z80;
                x86 (16-bit 8086);
                x86\_32 (IA-32);
                x86\_64 (AMD64)

                \item
                \textbf{\uline{Systems \& Platforms:}} \hspace{0pt}
                AIX;
                AmigaOS;
                Android;
                Apollo Domain OS;
                BSD (FreeBSD, OpenBSD, NetBSD);
                CP/M;
                DR-DOS;
                FreeRTOS;
                HPUX;
                IBM 360;
                iOS;
                Linux (kernel and distros);
                macOS;
                MINIX;
                MPE;
                MS-DOS;
                Multics;
                MTS;
                MVS;
                MyArc Geneve;
                NDOS;
                OS/2;
                QNX;
                RTOS;
                SGI IRIX;
                SGI ONYX;
                Solaris;
                TI-99/4A;
                Unix;
                VOS;
                VxWorks;
                Windows 3.0;
                Windows 3.1;
                Windows NT

                \item
                \textbf{\uline{Drivers \& Low-level Development:}} \hspace{0pt}
                Device Tree Overlays (Linux ARM platforms);
                HPUX drivers (68020, C);
                Linux kernel modules (loadable drivers, LKM);
                Linux VFS (Virtual Filesystem Switch);
                MPE drivers (Pascal);
                MS-DOS INT 13h BIOS-level block device handlers;
                macOS drivers;
                macOS I/O Kit (C++/Objective-C);
                low-level filesystem drivers;
                low-level filesystems - XFS and Clustered-XFS;
                PCI configuration space access;
                RSX-11 and VAX/VMS device drivers (MACRO-11, BLISS);
                SCSI-II Autochanger drivers;
                SCSI-II Magneto-optic drivers;
                Solaris Device Driver Interface (DDI/DKI);
                TI-99/4A CRU (Communications Register Unit) drivers (assembly);
                Windows 3.x DDK;
                Windows DDK (Device Driver Kit);
                Windows NT DDK;
                BIOS bootloader development;
                EFI (Extensible Firmware Interface) programming;
                GRUB bootloader configuration and customization;
                Secure Boot chain validation and integration.

                \item
                \textbf{\uline{Programming Languages \& Tools:}}  \hspace{0pt}
                .NET;
                6502 assembly;
                6809 assembly;
                8088 assembly;
                Alpha assembly;
                AMD bit-slice;
                AMD29K assembly;
                Apache Spark;
                Apollo DSEE (Domain Software Engineering Environment);
                ARM assembly;
                AWK;
                bash;
                bison;
                Boost (C++ Libraries);
                Borland C++;
                BoundsChecker;
                C;
                C\#;
                C++;
                C++ template metaprogramming;
                Carbon;
                ClearCase;
                COBOL;
                CodeWarrior;
                Cucumber;
                CVS;
                Eclipse (IDE);
                Enterprise Architect (Sparx Systems);
                flex;
                FORTH;
                FORTRAN;
                gcc;
                gdb;
                gem;
                Git (Version Control);
                Go (Golang);
                gprof;
                HyperCard;
                IDL (Interface Definition Language);
                IBM 360 assembly;
                IntelliJ IDEA (IDE);
                Itanium assembly;
                Java;
                JavaScript (browser);
                JCL (Job Control Language);
                JDK (Java Development Kit);
                JetBrains MPS;
                JNI (Java Native Interface);
                JProfiler;
                JUnit;
                ld;
                lex;
                Linux kernel modules;
                lldb;
                M68000 assembly;
                macOS I/O Kit;
                macOS SDK;
                make;
                MIPS assembly;
                Modern shell (sh, bash, zsh);
                MS-DOS INT 13h handlers;
                NetBeans (IDE);
                Node.js;
                npm;
                nvm;
                NUnit;
                OpenAPI (formerly Swagger);
                OpenGL;
                PA-RISC assembly;
                Pascal;
                Perforce (Version Control);
                Perl;
                PDP-11 assembly;
                Postman;
                PowerPC assembly;
                Process Explorer;
                Procmon;
                Protocol Buffers (protobuf);
                PsExec;
                Purify;
                Python;
                QTML;
                QuickDraw;
                QuickTime SDK;
                RCS (Version Control);
                RCA COSMAC 1802 assembly;
                React;
                RISC-V assembly;
                RSpec;
                RS6000 assembly;
                Ruby;
                Salad;
                Scala;
                SCCS (Version Control);
                sed;
                Selenium;
                shell (sh, csh);
                SoftICE;
                SPARC assembly;
                SQL (DDL and DML);
                Subversion (SVN Version Control);
                Swagger (OpenAPI predecessor);
                Sysinternals Suite;
                TI GPL (Graphics Programming Language);
                TMS9900 assembly;
                T-SQL;
                TypeScript;
                UML (Unified Modeling Language);
                Unix utilities (awk, sed, grep);
                Valgrind;
                VAX assembly;
                vi;
                vim;
                Visual Studio (IDE);
                VisualVM;
                VTune;
                webpack;
                Win32 API;
                Windows 3.x DDK;
                Windows DDK;
                Windows NT DDK;
                Windows SDK;
                Windbg;
                X11;
                Xcode;
                XNU kernel;
                Xt;
                x86 (16-bit 8086) assembly;
                x86\_32 (IA-32) assembly;
                x86\_64 (AMD64) assembly;
                yacc;
                yarn;
                Z80 assembly
                \item
                \textbf{\uline{UI \& Graphics:}} \hspace{0pt}
                AWT (Abstract Window Toolkit);
                Cross-Platform QuickTime;
                Direct2D;
                Fonts and Typography;
                Geometry;
                Glyphs;
                GTK+;
                HTML5 Canvas;
                Java2D;
                JavaFX;
                macOS UI;
                Microsoft GDI (Graphics Device Interface);
                OpenGL;
                Qt;
                Quartz 2D;
                QuickDraw 2D;
                QuickTime 1.x--3.x;
                SVG;
                Swing;
                Windows GUI;
                X11;
                \item
                \textbf{\uline{HID: Graphics, Video \& Sound:}} \hspace{0pt}
                DirectX;
                DisplayPort;
                dropping frames;
                FFmpeg;
                graphics performance;
                H.264;
                HDMI;
                jitter;
                latency;
                NVENC;
                OpenGL;
                sound card drivers;
                VGA drivers;
                Video Capture;
                Video Capture Card Drivers;
                Video Capture Cards;
                Video Compression;
                Video Decompression;
                video pipeline;
                video transcoding;
                WebRTC;
                YUV;
                \item
                \textbf{\uline{Statistical \& Optimization Methods:}} \hspace{0pt}
                6-sigma;
                A/B testing;
                gradient ascent;
                gradient boosting;
                gradient descent;
                K-means clustering;
                linear regression;
                logistic regression;
                monte carlo simulation;
                Multidimensional linear regression;
                simulated annealing;
                stochastic gradient descent;
                SVD (Singular Value Decomposition);
                \item
                \textbf{\uline{Networking \& Protocols:}} \hspace{0pt}
                802.11;
                BGP;
                Bluetooth;
                Bluetooth Low Energy;
                CAN bus;
                DNS;
                Ethernet;
                FTP;
                HTTP;
                HTTPS;
                IMAP;
                IPv4;
                IPv6;
                ISDN (Integrated Services Digital Network);
                leased lines;
                LoRaWAN;
                LU2;
                LU6.2;
                Modbus;
                MQTT;
                Novell NetWare;
                POP3;
                SMTP;
                SNA (System Network Architecture);
                SNMP;
                SSH;
                T1;
                TCP/IP;
                Telnet;
                TLS;
                Token Ring;
                UDP;
                WebSocket;
                Wi-Fi;

                \item
                \textbf{\uline{Enterprise \& Legacy Systems:}} \hspace{0pt}
                AS/400;
                CICS (Customer Information Control System);
                COBOL;
                DB2;
                HP 3000;
                JCL (Job Control Language);
                MPE;
                MTS (Michigan Terminal System);
                MVS (Multiple Virtual Storage);
                RSTS;
                RSX-11;
                System/360;
                System/370;
                TPF (Transaction Processing Facility);
                VAX/VMS;

                \item
                \textbf{\uline{Project \& Product Experience:}} \hspace{0pt}
                Apollo Computer Acquisition;
                Apple StarTrek;
                Beckman;
                BeyondNews;
                Bio-Rad;
                D. E. Shaw;
                EG\&G;
                Element Analytics;
                F5 Networks;
                Fisher Scientific;
                Forticom;
                GIMPS;
                Hewlett Packard;
                IdeaLab;
                Independa;
                Intacct;
                Intuit;
                NASA;
                NASDAQ;
                National Merit;
                Netflix Prize;
                Neustar;
                NYSE;
                Perkin-Elmer;
                Pixar;
                Tektronix;
                U.S. Coast Guard Auxiliary;
                ValueClick;
                Varian;
                Waters HPLC;

                \item
                \textbf{\uline{OS Internals \& Concepts:}} \hspace{0pt}
                Async distributed software;
                bootloaders;
                caching
                device driver abstraction;
                file descriptor management;
                Global-scale time synchronization algorithms;
                init systems;
                interrupt handlers;
                kernel scheduling;
                memory management;
                paging;
                preemption;
                process isolation;
                resource management;
                semaphores;
                system calls;
                virtual memory management;

                \item
                \textbf{\uline{Development Methodologies:}} \hspace{0pt}
                Agile;
                Behavior-Driven Development (BDD);
                Documentation-driven design;
                Kanban;
                Model-Driven Development (MDD);
                product management;
                Requirements before Design;
                Six Sigma;
                Scrum;
                Test-Driven Development (TDD);
                Toyota Production System (TPS);
                Waterfall;

                \item
                \textbf{\uline{Vintage Computing:}} \hspace{0pt}
                5" Floppy (FM/MFM formatting);
                8" Floppy (MFM formatting);
                Altair 8800;
                Apple II;
                Apple III;
                Commodore Amiga;
                Commodore PET;
                Commodore SuperPET;
                CP/M;
                CRT terminals;
                DEC PDP-11;
                IBM PC XT;
                KayPro;
                paper tape readers;
                S100 bus;
                Seequa Chameleon;
                Teletypes;
                TI-99/4A;
                TRS-80;

                \item
                \textbf{\uline{Language Support \& Internationalization:}} \hspace{0pt}
                bidirectional text;
                CJK (Chinese, Japanese, Korean) language support;
                codepage translation;
                collation algorithms;
                date and time formatting;
                font fallback;
                ICU (International Components for Unicode);
                ICU collation services;
                input method editors (IME);
                language negotiation;
                locale-aware string sorting;
                locale handling;
                LTR and RTL layout handling;
                multi-script rendering;
                natural language collation;
                number and currency formatting;
                pluralization rules;
                sort key generation;
                string normalization;
                surrogate pairs;
                text segmentation;
                transliteration;
                Unicode code points;
                Unicode Internals;
                Unicode normalization forms (NFC, NFD, NFKC, NFKD);
                Unicode scripts;
                UTF-16;
                UTF-8;
                variant selectors;
                zero-width joiners;

                \item
                \textbf{\uline{Human-Computer Interaction:}} \hspace{0pt}
                command-line interfaces;
                dialog systems;
                eye tracking;
                multimodal interfaces;
                Speaker-independent speech recognition;
                Speech synthesis;
                touchscreen UI;
                \item
                \textbf{\uline{Cloud Scale Platform Engineering:}} \hspace{0pt}
                ACME protocol (TLS automation);
                Alerting thresholds for metrics/logs;
                All test stages (unit, integration, e2e) integrated into CI pipelines;
                API backward compatibility;
                Apache Kafka;
                Apache Spark;
                Architecture Review Board (ARB) \& publishing cadence;
                Archive of CTO/Architecture policies;
                Attribute-based access control (ABAC);
                Audit log long-term retention;
                AuthN vendor policies;
                Blue/green deployments;
                Canary deployments;
                CCPA / GDPR privacy compliance;
                Ceph;
                Centralized log aggregation;
                Centralized policy engine (e.g., OPA);
                CI/CD enforcement with SAST, DAST, TDD;
                Cloud-init;
                Cloudflare (CDN);
                Code coverage gating;
                Consul;
                Container image SemVer tagging;
                Containerd;
                CoreDNS;
                CVE scanning policy;
                Data classification and handling policy;
                Data encryption at rest within storage and databases;
                Data encryption in transit between pods and nodes;
                Decision-Making Process (e.g., DACI);
                Declarative orchestration (e.g., Kubernetes);
                Delegated permissions and admin domains;
                DDoS protection policies;
                Distributed tracing (OpenTelemetry);
                DNS vendors and zone management policies;
                Docker;
                Domain registrar vendors and policies;
                ElasticSearch;
                etcd;
                Feature flag routing;
                Federated authentication (OIDC/SAML);
                Fine-grained RBAC;
                Flannel;
                Git commit SHA tagging;
                GitOps change management;
                gRPC;
                Health/liveness/readiness probes;
                Helm;
                Honey Pots;
                Horizontal autoscaling policies;
                Hyperscalers (AWS, GCP, AZURE);
                IDS (Intrusion Detection);
                Immutable audit trails;
                Integration test harness;
                IPS (Intrusion Prevention);
                ISO 27001 / NIST optional benchmarks;
                Istio;
                JWT or SPIFFE/SPIRE identity;
                Kubernetes (K8s);
                Linkerd;
                Load and chaos testing;
                Log redaction and privacy rules;
                Log retention policy;
                Logging access controls;
                Logging format: structured JSON;
                Max payload sizes (upload/request/response);
                Metric ontologies (RED, USE);
                Metrics vendor(s) and retention policy;
                MinIO;
                Minimum unit test coverage threshold (e.g., 90\%);
                mTLS enforcement (Istio, Linkerd);
                Multiple entry points per container allowed;
                Multi-zone / HA deployment policy;
                Network load shaping to gradually shift traffic;
                NGINX;
                One process per container;
                Open Policy Agent (OPA);
                OpenTelemetry;
                OWASP ruleset enforcement (blocking mode only);
                Policy versioning and rollback;
                PostgreSQL;
                Production-like staging environments;
                Prometheus;
                Pulumi;
                Rate limits – global and per-service;
                Redis;
                Regression test suite on merge;
                Registry policies: promotion lifecycle (dev → stage → prod);
                Regular rotation and management of encryption keys: SLA of 5 minutes or less for entire network; Services reload keys upon receiving a specific signal;
                Required DNS records (A, AAAA, CAA, CNAME, NS, PTR, SPF, DKIM, DMARC, MTA-STS, MX etc.);
                Required IPv4 and IPv6 support;
                Required log fields (timestamp, trace\_id, etc.);
                Required security headers (default values);
                Resource quotas and limits;
                RFC Process for architecture decisions;
                Rollback plan required for changes;
                S3;
                SBOM (Software Bill of Materials);
                Schema migration strategy;
                SDK-level log enforcement;
                Secrets management;
                Secure Development Lifecycle (SDL);
                Secure key vault integration;
                Service-level AuthZ routing;
                Service mesh enforcement;
                Session management policies;
                Sharding;
                SLA-based alerting and load shaping;
                SLO dashboards;
                SOC-2 Type II compliance;
                Structured audit logging;
                Synthetic functional tests in production;
                Terraform;
                TLS cipher suite and protocol policy;
                Token lifecycle and revocation;
                Trace propagation and sampling;
                Transactional or rollback-safe deploys;
                Unified taxonomy \& namespace schema;
                Vault (HashiCorp);
                Versioned architecture standards;
                Wildcard \& internal DNS policies;
                Zookeeper;
                Zero-downtime deployments;

                \item
                \textbf{\uline{IoT edge to cloud:}} \hspace{0pt}
                Access Control (IAM);
                Analytics Services;
                API Gateway;
                Asset Models / Digital Twins;
                Azure Arc;
                Azure Data Explorer (ADX);
                Azure Digital Twins (ADT);
                Azure IoT Central;
                Azure IoT Edge;
                Azure IoT Hub;
                Azure Stream Analytics (ASA);
                Azure Time Series Insights (TSI);
                AWS CloudFormation;
                AWS Greengrass Core;
                AWS Greengrass Lambda Functions;
                AWS IoT Analytics;
                AWS IoT Core;
                AWS IoT Device Defender;
                AWS IoT Device Management;
                AWS IoT Events;
                AWS IoT FleetWise;
                AWS IoT SiteWise;
                AWS Lambda;
                BACNET;
                Battery Management (edge);
                Data Brokers (MQTT / AMQP / HTTP);
                Device Firmware / RTOS;
                Device Health Monitoring;
                Device Identity Registry;
                Device Provisioning Service (DPS);
                Digital Twin Modeling Language (DTDL);
                Edge Gateway / Industrial PC;
                Edge Impulse;
                Edge ML Frameworks (TensorFlow Lite);
                Edge Runtime (Azure IoT Edge / Greengrass Core);
                Factory Keyed (Zero Trust);
                Firmware Over-the-Air (FOTA);
                Function-as-a-Service (FaaS);
                gRPC / REST APIs;
                Hardware Security Module (HSM / TPM);
                Ingestion Pipelines;
                KubeEdge;
                Local Message Broker (e.g., Mosquitto, Greengrass Broker);
                Local ML Inference;
                Logging and Metrics Collection;
                Mesh Wireless (BLE, WiFi);
                Message Queues (MQTT, Kafka, NATS, RabbitMQ);
                Microcontroller / SoC;
                Network Protocols (MQTT, CoAP, HTTP);
                Over-the-Air Updates (OTA);
                PLC / SCADA Integration;
                Remote Management Tools;
                Rule Engines;
                Security Policies (IAM, RBAC);
                Sensor Drivers;
                Sensor Modules;
                Stream Processing Frameworks;
                Telemetry Collection;
                Telemetry Routing;
                Time Series Database (TSDB);
                TLS Mutual Authentication;
                Twin Synchronization;
                Visualization Dashboards;
                Workload Orchestration (Kubernetes, ECS, AKS);
            \end{itemize}
        }
    \end{multicols}

\end{document}
